Assessing the Accessibility of Protein Tunnels using Motion Planning

### 2017-12-08 paper.tex
missing files

\author{Vojt\v ech Von\' asek$^{1}$,
    Barbora Kozl\'\i kov\'a$^{2}$,
    Adam Jur\v{c}\'\i k$^{2}$,
    Ond\v{r}ej V\'{a}vra$^{3}$,
    Martin Saska$^{1}$
\thanks{$^{1}$ Faculty of Electrical Engineering,  Czech Technical University in Prague, 
Technick\'a 2, 166 27 Prague, Czech Republic, {\tt vonasek@labe.felk.cvut.cz};
$^{2}$ Faculty of Informatics, Masaryk University, Botanick\'a 68a, 602~00 Brno, Czech Republic;
$^{3}$ Loschmidt Laboratories, Department of Experimental Biology RECETOX, Faculty of Science, Masaryk University
    Kamenice 5, 625 00 Brno, Czech Republic}
    }
%The presented work has been supported by the Czech Science Foundation (GA{\v C}R) under research project No. 17-07690S.
%Access to computing and storage facilities owned by parties and projects contributing to the National Grid Infrastructure MetaCentrum, provided under the programme "Projects of Large Infrastructure for Research, Development, and Innovations" (LM2010005), is greatly appreciated.
%}%
%}
%}

@@

The traversability rate is shown in Tab.~\ref{tab::main}. 
The table shows the traversability rate only for the main tunnel, as no trajectory was found in the side tunnel for most of the tested ligands.
The only exception is the ligand m037t, which was the smallest tested one (with 8 atoms), and therefore it could fit even inside the side tunnel.
The traversability rate is higher for the most scaled-down ligands ($\smin=0.5$) and it decreases with the increasing scale.
For the ligand DCP, trajectories were detected in 10~\% of frames for scale $\smin=0.8$ and the traversability of this ligand
to the active site was approved by a separate MD simulation~\cite{marques2017catalytic}.
%For the scale $\smin=0.8$, the most traversable frames were detected for ligands DCP (18~\%), m037t (45~\%) and m038t (30~\%).
%The traversability of the DCP ligand to active site was approved by a separate MD simulation~\cite{marques2017catalytic}.
Based on the number of success rate, also other ligands (e.g., m037t with trav. rate 45~\% at $\smin$=0.8) seem to be promissing.
However, their binding was not approved yet by MD dedicated simulations.

@@

Judging the accessibility of the active site by a ligand based only on the traversability rates may be problematic if the ligand is scaled too much.
Instead of using a single version of the scaled-down ligand, we only limit the minimal scale used by the planner (the parameter $\smin$) and
the planner prefers to apply the large scales in possible.
%The expansion step always prefers to expand the tree by a collision-free configuration with the maximal scale.
%that leads to a new collision-free configuration.
The information about the scales used at the trajectory points (which can be extracted from the configuration tree of the proposed planner) brings the biochemists new knowledge about the tunnels, which was impossible to get using the previous solutions.
Our proposed visualization allows the biochemists to observe those parts of the tunnel where the ligand had to be scaled-down significantly (red parts of trajectories in Fig.~\ref{fig::dcp}), and parts where the ligand could be enlarged again (green parts in Fig.~\ref{fig::dcp}).
According to the results of the motion planing, the DCP needs to shrink down to ca. 75~\% of its size (Fig.~\ref{fig::dcp}, point B).
However, as both the protein and ligand are mutually influence each other, it is possible that even a very narrow part of the tunnel is traversable.
This was observed in the MD simulation of the DCP ligand in the 1CQW protein, which revealed that DCP can pass the main tunnel~\cite{marques2017catalytic}.

>>>>

\author{Vojt\v ech Von\' asek$^{1}$,
    Barbora Kozl\'\i kov\'a$^{2}$,
    Adam Jur\v{c}\'\i k$^{2}$,
    Ond\v{r}ej V\'{a}vra$^{3}$,
    Martin Saska$^{1}$
\thanks{$^{1}$ Faculty of Electrical Engineering,  Czech Technical University in Prague, 
Technick\'a 2, 166 27 Prague, Czech Republic, {\tt vonasek@labe.felk.cvut.cz};
$^{2}$ Faculty of Informatics, Masaryk University, Botanick\'a 68a, 602~00 Brno, Czech Republic;
$^{3}$ Loschmidt Laboratories, Department of Experimental Biology RECETOX, Faculty of Science, Masaryk University
    Kamenice 5, 625 00 Brno, Czech Republic. 
The presented work has been supported by the Czech Science Foundation (GA{\v C}R) under research project No. 17-07690S.
}}
%The presented work has been supported by the Czech Science Foundation (GA{\v C}R) under research project No. 17-07690S.
%Access to computing and storage facilities owned by parties and projects contributing to the National Grid Infrastructure MetaCentrum, provided under the programme "Projects of Large Infrastructure for Research, Development, and Innovations" (LM2010005), is greatly appreciated.
%}%
%}
%}

@@

The traversability rate is shown in Tab.~\ref{tab::main}. 
The table shows the traversability rate only for the main tunnel, as no trajectory was found in the side tunnel for most of the tested ligands.
The only exception is the ligand m037t, which was the smallest tested one (with 8 atoms), and therefore it could fit even inside the side tunnel.
The traversability rate is higher for the most scaled-down ligands ($\smin=0.5$) and it decreases with the increasing scale.
For the ligand DCP, trajectories were detected in 10~\% of frames for scale $\smin=0.8$ and the traversability of this ligand
to the active site was approved by a separate MD simulation~\cite{marques2017catalytic}.
%For the scale $\smin=0.8$, the most traversable frames were detected for ligands DCP (18~\%), m037t (45~\%) and m038t (30~\%).
%The traversability of the DCP ligand to active site was approved by a separate MD simulation~\cite{marques2017catalytic}.
Based on the number of success rate, also other ligands (e.g., m037t with traversability rate 45~\% at $\smin$=0.8) seem to be promissing.
However, their binding was not approved yet by MD dedicated simulations.

@@

Judging the accessibility of the active site by a ligand based only on the traversability rates may be problematic if the ligand is scaled down too much.
Instead of using a single version of the scaled-down ligand, we only limit the minimal scale used by the planner (the parameter $\smin$) and
the planner prefers to apply the higher scales if possible.
%The expansion step always prefers to expand the tree by a collision-free configuration with the maximal scale.
%that leads to a new collision-free configuration.
The information about the scales used at the trajectory points % (which can be extracted from the configuration tree of the proposed planner) 
brings the biochemists new knowledge about the tunnels, which was impossible to get using the VD-based computations.
The  visualization allows the biochemists to observe those parts of the tunnel where the ligand had to be scaled-down significantly (red parts of trajectories in Fig.~\ref{fig::dcp}), and parts where the ligand could be enlarged again (green parts in Fig.~\ref{fig::dcp}).
According to the results of the motion planing, the DCP needs to shrink down to ca. 75~\% of its size (Fig.~\ref{fig::dcp}, point B).
However, as both the protein and ligand are mutually influence each other, it is possible that even a very narrow part of the tunnel is traversable.
This was observed in the MD simulation of the DCP ligand in the 1CQW protein, which revealed that DCP can pass the main tunnel~\cite{marques2017catalytic}.
